---
title: "Ag Machine Pass Numbering"
author: "Eric Coronel"
date: "August 6, 2018"
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose of this script

Data collected by agricultural equipment is not user friendly, especially when within-field data manipulation is needed. It is simple to calculate average yield or fuel usage per field. The challenge is obtaining attributes from individual field passes inside the field when there are no identifiers to perform subsettings or groupings.  
The code included here uses time difference in seconds between observations to separate machinery passes. The threshold for separating passes can be modified.

## Packages needed
```{r packages, eval = T, message = F, warning = F}

library(dplyr)
library(ggplot2)
library(rgdal)

```

## Importing the seeding shapefile

A publicly available seeding shapefile can be found at a [Deere website](https://developer.deere.com/#!documentation&doc=.%2Fmyjohndeere%2FfieldOperations.htm&anchor=). The sample Seeding shapefile can be downloaded from the following [link](https://developer.deere.com/content/documentation/merriweather_seeding.zip). The zip file needs to be decompressed into a temp file before importing. 

```{r import, eval = T}
# Download zipfile into tempfile
url <- "https://developer.deere.com/content/documentation/merriweather_seeding.zip"
tmpfn <- tempfile()
zipfile <- download.file(url, tmpfn)
files <- unzip(tmpfn, exdir = tempdir())
```

Once decompressed it can be imported using `rgdal::readORG` with `pointDropZ = TRUE` argument because the shapefile has an empty Z column; likely an error during the creation of the file. I created a working copy to avoid manipulating the original file since it takes about 30 seconds to import the original shapefile.

```{r}
# Importing the shapefile
seeding_shapefile <- readOGR(dsn = files[5], layer = "Merriweather Farms-JT-01-Corn", pointDropZ = TRUE)

# Working copy
work_copy <- seeding_shapefile
```

## First processing

This chunk does the following:

* Creates a datetime variable.
* Calculates time difference in seconds among consecutive records.
* Creates an indicator variable that assigns 0 if time difference is between 0 to 6 seconds, and 1 to all others.
* Calculates the cumulative sum of the indicator variable.   

Since data are arranged in ascending time order, the cumulative sum of the indicator variable gives us the total number of passes.

```{r processing_a, eval = T}

work_copy@data <- work_copy@data %>%
  mutate(datetime = as.POSIXct(Time, format = "%m/%d/%Y %H:%M:%S"),
         diff1 = c(NA, diff(datetime)),
         indicator = ifelse(diff1 %in% c(0:6), 0, 1),
         pass_num = cumsum(indicator))

```

## Calculating observations per pass

Another useful attribute for grouping or subsetting is knowing the number of observations per pass. This code also creates a random order of pass numbers (`pass_scramble`) which is useful for later visualization.

We first create a new file, `obs_pass`, with the variables mentioned above. 

```{r obs_per_pass, eval = T}

obs_pass <- work_copy@data %>%
  group_by(pass_num) %>%
  summarize(obs_per_pass = n()) %>%
  mutate(pass_scramble = factor(sample(pass_num)))

```

## Joining new data to seeding shapefile

This chunk:

* Joins the `obs_pass` file to the seeding shapefile.
* Removes `datetime`, `diff1`, and `indicator` variables.
* Makes actual pass numbers variable, `pass_num`, a factor.

```{r joining, eval = T}

work_copy@data <- work_copy@data %>%
  left_join(obs_pass, by = "pass_num") %>%
  select(-datetime, -diff1, -indicator) %>%
  mutate(pass_factor = factor(pass_num))

```

## Giving the coordinates for each point

It is often useful to have the coordinates as part of the attribute table.

```{r coords, eval = T}

# Creating file with coordinates
xycoords <- work_copy %>%
  coordinates() %>%
  data.frame

# Joining the coordinates with the shapefile
work_copy@data <- work_copy@data %>%
  bind_cols(xycoords)

```

## Ploting final version

The variable `pass_scramble` is used for coloring to help with visualization.

```{r plotting, eval = T}

# Using a subset for faster display
ggplot(filter(work_copy@data, pass_num %in% 20:25), 
       aes(x = coords.x1, y = coords.x2, color = pass_scramble)) +
  geom_point() +
  scale_color_hue(h = c(0, 360), c = 100, l = c(30, 60)) +
  labs(x = "Longitude",
       y = "Latitude",
       color = "Passes \n (order was randomized)")

# Plotting the entire shapefile
ggplot(work_copy@data, aes(x = coords.x1, y = coords.x2, color = pass_scramble)) +
  geom_point() +
  scale_color_hue(h = c(0, 360), c = 100, l = c(30, 60)) +
  labs(x = "Longitude",
       y = "Latitude",
       color = "Passes \n (order was randomized)")

```